---
title: "Reimagining the ATP Tour"
output: html_document
author: "Joshua Snoke"
date: 01-13-2024
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(geosphere)
library(TSP)
library("ggplot2")
theme_set(theme_bw())
library("sf")

library("rnaturalearth")
library("rnaturalearthdata")

world = ne_countries(scale = "medium", returnclass = "sf")

here::i_am('atp_optimal.R')
atp_data = read_csv('atp_optimal.csv') %>%
    filter(tournament %in% c("australian open",
                             "shangai",
                             "tokyo",
                             "indian wells",
                             "acapulco",
                             "miami",
                             "washington",
                             "cincinnati",
                             "montreal",
                             "us open",
                             "queens",
                             "wimbledon",
                             "monte carlo",
                             "rome",
                             "barcelona",
                             "madrid",
                             "roland garros",
                             "paris",
                             "vienna",
                             "turin")) %>%
    select(-simul)
```

## The Current Tour, Crisscrossing the Globe

The current ATP tour crisscrosses the globe numerous times, expending incredible amounts of jet fuel. As an avid tennis fan, I have often wonder how much this could be reduced with a re-imagined tour. The problem is finding the shortest path between a set of points is simple enough (https://en.wikipedia.org/wiki/Travelling_salesman_problem), so let's see what happens when we through the current tour at it.

First, a few notes and caveats.

(1) I use the ATP tour, but an equally important assessment could look at the WTA tour. I decided to only do one because this makes it simpler, and I use the ATP because currently the tournament schedule seems a bit more consistent (e.g., no uncertainty over tour finals, etc.). And hey, if we are going to reimagine the tour, why not imagine a world where all these tournaments are combined women and men. What a awesome world that would be.
(2) I limited to the tour to 20 events, in keeping with the ATP ranking system. Of course, incredible amounts of jet fuel are used to fly around to 250s all over the place, but in my imagination we could organize 250s in the same or neighboring cities (or cities along the path) to the main 20 events. Or! Maybe tennis could shrink its calendar and focus its effort on making a really good 20 event season, with space for team events in the rest of the calendar. We are dreaming here after all.

### Mapping the Current Tour

I kept the four majors, tour finals, nine 1000s, and six of the 500s (I am sure some people will be mad which ones I chose but you could easily swap a few). For location, I collected the longitude and latitudes of the nearest major airport for each tournament city, resulting the following list.

```{r atp, echo=FALSE}
kable(atp_data)
```

```{r, echo=FALSE}
dist_all = function(targets, fixed){
    distm(c(atp_data$lon[targets], atp_data$lat[targets]), 
          c(atp_data$lon[fixed], atp_data$lat[fixed]), 
          fun = distVincentyEllipsoid)
}

dist_wrapper = function(fixed, input_data){
    sapply(c(1:nrow(input_data)), 
           dist_all,
           fixed)
}

dist_matrix = sapply(1:nrow(atp_data),
                     dist_wrapper,
                     atp_data)
rownames(dist_matrix) = colnames(dist_matrix) = c(atp_data$tournament)

## add dummy to force Turin as end
## ## set inf distance from aus to dummy but zero in other way
dist_matrix = bind_cols(dist_matrix, 
                        'dummy' = c(Inf, rep(Inf, nrow(dist_matrix) - 2), 0))
dist_matrix[(nrow(dist_matrix) + 1), ] = as.list(c(0, rep(Inf, nrow(dist_matrix) - 2), 0, 0))

dist_matrix = as.matrix(dist_matrix)
## add names
rownames(dist_matrix) = colnames(dist_matrix) = c(atp_data$tournament, 'dummy')

## current 
atsp = ATSP(dist_matrix)
tour_current = solve_TSP(atsp, 
                         method = 'identity',
                         start = 1)

## optimal
suppressWarnings({tour_optimal = solve_TSP(atsp, 
                                           method = 'nearest_insertion',
                                           start = 1,
                                           repetitions = 100)})

## reorganized
city_manual = c("australian open",
                "shangai",
                "tokyo",
                "indian wells",
                "acapulco",
                "miami",
                "washington",
                "cincinnati",
                "montreal",
                "us open",
                "queens",
                "wimbledon",
                "monte carlo",
                "rome",
                "barcelona",
                "madrid",
                "roland garros",
                "paris",
                "vienna",
                "turin",
                "dummy")

atsp_manual = ATSP(as.matrix(dist_matrix[city_manual, 
                                         city_manual]))
tour_manual = solve_TSP(atsp_manual, 
                         method = 'identity',
                         start = 1)
```

The current tour has a lot of crisscrossing! Turns out if you flew along the current path to all 20 tournaments, that would be roughly 62,667 kilometers! That is equal to flying around the whole earth more than one and a half times...

```{r, echo=FALSE, fig.width=12}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = atp_data, aes(y = lat, x = lon), color = 'black') +
    geom_segment(data = atp_data,
                 aes(x = lon, 
                     y = lat, 
                     xend = lead(lon, 1, default = 7.647438),
                     yend = lead(lat, 1, default = 45.19612)),
                 arrow = arrow(length = unit(0.15, "cm")),
                 color = '#e41a1c')
```

### Finding a Shorter Path Between Tournaments

Now, let's see what happens if we find the shortest path that goes through all 20 tournaments. Using a pretty simple algorithm we find the following optimal path:

```{r, echo=FALSE}
kable(tibble(`shortest path` = labels(tour_optimal)) %>%
        filter(`shortest path` != 'dummy') %>%
        left_join(atp_data, by = c('shortest path' = 'tournament')))
```

We still have to go pretty far, but now it's only 37,605 kilometers. We reduced the distance flown (and emissions produced) by 40%. 

```{r, echo=FALSE, fig.width=12}

optimal_data = tibble(`shortest path` = labels(tour_optimal)) %>%
    left_join(atp_data, by = c('shortest path' = 'tournament')) %>%
  filter(`shortest path` != 'dummy')

ggplot(data = world) +
    geom_sf() +
    geom_point(data = optimal_data %>% na.omit, aes(y = lat, x = lon), color = 'black') +
    geom_segment(data = optimal_data %>% na.omit,
                 aes(x = lon, 
                     y = lat, 
                     xend = lead(lon, 1, default = 7.647438),
                     yend = lead(lat, 1, default = 45.19612)),
                 arrow = arrow(length = unit(0.15, "cm")),
                 color = '#377eb8')
```

### Creating a Shorter, Sensible Tour

Now, this tour does not make total sense even if it is the shortest path, so let's rearrange a few things. Let's ensure that we have the tournaments arranged by surface and that each swing ends in the appropriate major. 

(1) Starting in Australia, we have a hardcourt season that goes through Asia and then North America, culminating in the U.S Open. This would like be early-mid summer.
(2) Next, we go to Europe and have a short grass court swing and Wimbledon.
(3) Still in Europe we move to clay for the late-summer/early fall ending in Roland Garros.
(4) Finally, we end the year with some indoor tournaments and the tour ending finals.

Hey, that sounds pretty good. Here's what we have:

```{r, echo=FALSE}
kable(tibble(`ideal path` = labels(tour_manual)) %>%
        filter(`ideal path` != 'dummy') %>%
        left_join(atp_data, by = c('ideal path' = 'tournament')))
```

Having moved a few tournaments around to get the majors at the end of each swing, we have to fly 38,555 kilometers. That's not a whole lot more than the shortest path. And still a reduction of 38% in km/emissions from the current tour!

```{r, echo=FALSE, fig.width=12}

ideal_data = tibble(`ideal path` = labels(tour_manual)) %>%
    left_join(atp_data, by = c('ideal path' = 'tournament')) %>%
  filter(`ideal path` != 'dummy')

ggplot(data = world) +
    geom_sf() +
    geom_point(data = ideal_data %>% na.omit, aes(y = lat, x = lon), color = 'black') +
    geom_segment(data = ideal_data %>% na.omit,
                 aes(x = lon, 
                     y = lat, 
                     xend = lead(lon, 1, default = 7.647438),
                     yend = lead(lat, 1, default = 45.19612)),
                 arrow = arrow(length = unit(0.15, "cm")),
                 color = '#4daf4a')
```

### Final Thoughts

This little excercise was not rocket science, but that is possibly the most important point. The ATP tour (and tennis in general) simply does not need to cross the globe as much as it does. Without changing any of the main tournaments, we could reduce the distance flown by 38%(!) simply by rearranging the order. This is what they call "low hanging fruit."

I'm sure someone could tweak this, and I'm interested to hear comments or possible improvements. But at the end of the day, even I was amazed to see just how easily you could reduce the waste of flying around the globe by reordering the tour, while keeping its fundamental structure the same.

```{r, echo=FALSE}
tibble(`current path` = labels(tour_current),
       `shortest path` = labels(tour_optimal),
       `ideal path` = labels(tour_manual)) %>%
    filter(`current path` != 'dummy') %>%
    kable
```


```{r, echo=FALSE,include=FALSE}
tour_length(tour_current)
labels(tour_current)

tour_length(tour_current)
labels(tour_current)

tour_length(tour_optimal)
labels(tour_optimal)

tour_length(tour_manual)
labels(tour_manual)

## fraction
tour_length(tour_optimal) / tour_length(tour_current)
tour_length(tour_manual) / tour_length(tour_current)
tour_length(tour_optimal) / tour_length(tour_manual)
```





